Mobile Assets in Semantic Digital Twins

### 2023-05-20 paper/main.tex
Add backup of paper

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{comment}
\usepackage{duomasterforside}
\usepackage{listings}
\usepackage{float}
\usepackage{fancyvrb}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{fvextra}
\usepackage{soul}
\usepackage{minted}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{multirow}
\graphicspath{ {./graphics/} }
\usepackage[backend=biber,natbib,style=numeric]{biblatex}
\usepackage{hyperref}
\addbibresource{references.bib}

@@

\section*{Abstract}
This thesis proposes a solution to the challenge of handling mobile assets in a semantic digital twin based on a dynamic asset model in which static data (existing building infrastructure) is separated from dynamic data (smartphone's physical location). The digital twin (DT) is created in Semantic Micro Object Language (SMOL) for easy integration with semantic technologies, such as semantic knowledge graphs. In order to semantically enrich the DT further, a model of a building is created in Web Ontology Language (OWL). A hybrid smartphone application is implemented having the mobile assets in mind, which sends their updated status. The privacy of the data subjects is protected as much as possible. The DT is an up-to-date digital representation of the PT, and it reconfigures in near real-time if the PT or any of the mobile assets inside it changes. If a person is located inside a dangerous area, the app warns the appropriate user from the informed decision sent by the DT.

@@

Handling mobile assets in a DT based on an \emph{asset model} (i.e. a file that contains an organized description of an asset's composition and properties \cite{kamburjan_twinning-by-construction_2022}) that is \emph{dynamic} is unexplored territory.

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.14]{graphics/initial_thesis_overview.png}
    \caption{Overview of \emph{main} architectural components in this thesis}
    \label{fig:initial_components}
\end{figure}

@@

\subsubsection{For the scope of this thesis}
As of today, a DT is referred to as a digital representation of some physical system \cite{grieves_digital_2017}, that has no restriction on the size or scale of it, which means the entity could be a single asset or a system of systems for that matter \cite{li_digital_2022, waszak_let_2022}. It exists in real life and is commonly referred to as the physical twin (PT), and is observed by the DT in near real-time. There should be a bidirectional flow of data between the DT and the PT throughout the life cycle of the PT. In other words, the DT gets updated sensor data from the PT, and sends informed decisions back to it \cite{madni_leveraging_2019, waszak_let_2022, kamburjan_digital_2022}. From this, and for the scope of this thesis, a DT is a digital representation of a system that tracks many mobile assets. The DT sends informed decisions to the physical counterpart in which the appropriate mobile assets perform actions.

There are multiple misconceptions about DTs. One such misconception is that a digital shadow (DS) is some sort of a DT. This misconception often leads people to build shadows instead of DTs. \cite{fuller_digital_2020, li_digital_2022}.  Although very similar, a digital shadow (DS) does not have the bidirectional data flow that a DT does. Instead, there is a one-way flow from the PT to the digital counterpart, which often leads to a change in the DS, but not vice versa \cite{kritzinger_digital_2018, li_digital_2022}. Secondly, as told by \citeauthor{fuller_digital_2020}, a common misconception is that the DT must be an exact 3D copy of a physical entity, such as a building.

@@

To sum up, it is important to understand what a DT is and what it is not. We should try to "close the loop" by sending informed decisions back to the physical counterpart so that it reflects the newest state of the DT. A DS could be used alongside the DT with the only goal of being a timestamped snapshot of some specific physical asset \cite{bergs_concept_2021}. When implementing one should also think about the challenges of digital twins, such as privacy concerns and costs. We should not get too caught up with modeling an exact 3D replica of an asset. On the other hand, we \emph{should} use existing semantic technologies such as knowledge graphs for formal knowledge representation from a variety of data sources.

@@

OWL documents can easily be created (from export), loaded into, and maintained in ontology editors, such as Protégé\footnote{\url{https://protege.stanford.edu}}. Such a tool can be used to give semantics to e.g. a model of a building, as well as spending less time on creating and maintaining it compared to BIM.

@@

\begin{figure}[H]
    \centering
    \caption{A simple SMOL program}
    \label{fig:smol_program}
    \begin{Verbatim}[frame=single]
class Smartphone(Int id, String name, Boolean isBlackBerry) end

@@

\subsection{Reasoning}\label{subsec:Reasoning}
OWL was previously mentioned in Section \ref{subsubsec:Technologies} as a language for expressing rich knowledge of concepts and relations between entities. We should keep in mind, however, that this formalized knowledge should be reasoned over and put in context to draw conclusions based on criteria. There are many technologies that let us do this.
\subsubsection{HermiT}
HermiT is a reasoner for OWL 2 \cite{glimm_hermit_2014}, which is an upgraded version of OWL. \citeauthor{glimm_hermit_2014} describes that the reasoner support both object and data property classification, as well as SPARQL query answering. HermiT is built into the ontology editor Protégé and can be used to reason over an ontology directly in the editor. Other plugins for reasoning can be added to the ontology editor, but it is convenient that HermiT comes with the installation.

@@

There also exist other technologies that provide reasoner tools. Some notable mentions include, but are not limited to:
\begin{itemize}
    \item{\textbf{Apache Jena}} is an open source framework that has the Inference API which include reasoner tools that can be used to reason over data, as well as checking content of triple stores \cite{noauthor_apache_nodate-1}. The source code is publicly available\footnote{\url{https://github.com/apache/jena}}.
    \item{\textbf{OWL API}} is a Java framework that lets us create, serialize or manipulate ontologies written in OWL \cite{noauthor_owl_nodate}, and its source code is openly accessible\footnote{\url{https://github.com/owlcs/owlapi}}. Its recent development is more focused towards OWL 2 however.
    \item\hyperref[subsec:SMOL]{\textbf{SMOL}} provide reasoning tools through its interpreter, which implements \emph{semantic lifting} (i.e. generating a knowledge graph from the current state of a SMOL program) \cite{noauthor_semantic_nodate}.

@@

\subsection{Research questions (RQ1, RQ2, RQ3)}
\subsubsection{RQ1}\label{subsubsec:RQ1}
RQ1 is concerned with creating a dynamic asset model of a simple building. From the descriptions of asset models in Section \ref{subsec:AssetModel}, an ontology of a simple building was created, as shown in Figure \ref{fig:building_ontology}. This asset model is a simplified version of an exported OWL document (file). It can be created in an ontology editor, such as Protégé. Although other ontology editors can be used, we should use Protégé due to its user-friendliness and the extensive documentation provided in its user guide \cite{horridge_practical_2011}. This exported file is provided to and used by the digital twin (DT) by using the SMOL interpreter.

@@

\begin{Verbatim}[commandchars=\\\{\}, breakanywhere=true]
### \hl{http://www.semanticweb.org/oscarlr/ontologies/2023/2/building#smartphone5}
\hl{:smartphone5} rdf:type owl:NamedIndividual ,
                      :MovableEntity ;
             :movableEntityId "\hl{5}" .
\end{Verbatim}

Section \ref{subsubsec:DynamicAssetModel} further examines what a \emph{dynamic} asset model entails.

@@

\subsubsection{RQ3}
RQ3 is concerned with how we can "separate dynamic data (smartphone’s physical location) from static data
(building infrastructure) in the asset model". Hypothesis (H) states that the asset model should be of a simple building. The rooms with a square area constitute it. In contrast to building infrastructure, mobile assets that are tracked can change at any time, as well as their continuously changing physical locations. As described in Section \ref{subsec:Scope}, we are not concerned with other movable entities in the building domain apart from the mobile assets. And, even though the data of the building infrastructure may change, it will not change during program execution in SMOL. This is not the case with the time series data accessed from InfluxDB in SMOL. However, one way to separate the dynamic data from the static data is by examining \emph{how} often changes occur through time, where separate means not affecting one another. More specifically, when reloading the asset model, the static data should not be affected by the data data considered dynamic. Further analysis of static and dynamic data is described in Section \ref{subsubsec:StaticAndDynamicData}, and also later discussed in Section \ref{subsec:DiscussDynamicData}.

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.16]{graphics/static_built_environment.png}
    \caption{A UML class diagram that illustrates the separation of static and dynamic behavior}
    \label{fig:static_built_environment}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.22]{graphics/static_dynamic_asset_model.png}
    \caption{The separation of \textbf{Static} and \textbf{Dynamic} data, illustrated in blue and red, respectively}
    \label{fig:static_dynamic_asset_model}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.12]{graphics/wireframe_app.png}
    \caption{Shows a wireframe of the UI in the app}
    \label{fig:wireframe_app}
\end{figure}

@@

Then we look at the final version in Figure \ref{fig:components}, and describe what was implemented. For each of the larger architectural components in the figure, we reference the section that describes it in Chapter \ref{sec:Background}, and the design decision that enables it in Chapter \ref{sec:Analysis}. We also reference the corresponding \emph{item} in the structural requirement \emph{part} we are in.

@@

An asset model (ontology) of a simple building was also created, as shown in Figure \ref{fig:simple_building} (this time with coordinates as data properties which were "added" to individuals of the \verb|MovableEntity| class suggested in Section \ref{subsubsec:DynamicAssetModel}. We used Protégé to create the domain and exported it as an OWL document (file). This file was then accessed in a new SMOL program based on the background information in Section \ref{subsubsec:Technologies}, Section \ref{subsubsec:SMOL}, and from further analysis in Section \ref{subsubsec:StaticAndDynamicData}.

@@

Figure \ref{fig:server_forward_informed_decision} shows how the informed decision from the DT is forwarded to Firebase.
\begin{figure}
    \centering
    \begin{small}
    \begin{minted}[breaklines]{java}
private void warnEndangeredClients(Map<String, String> results) {
    reference = FirebaseDatabase.getInstance().getReference("endangered");
    DatabaseReference.CompletionListener completionListener = (databaseError, databaseReference) ->
            System.out.println("Forwarded informed decision to PT!");
    reference.setValue(results, completionListener);
    reference.removeValue(completionListener);
}
    \end{minted}
    \end{small}
    \caption{Forward informed decision to PT through Firebase}
    \label{fig:server_forward_informed_decision}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minted}{java}
public void insertDataPoint(Object object) {
    WriteApiBlocking writeApi = db.getWriteApiBlocking();
    writeApi.writeMeasurement(bucket, org, WritePrecision.MS, object);
}
    \end{minted}
    \caption{Inserting a measurement point (timestamped status message) into the Data-bucket in InfluxDB}
    \label{fig:insert_point}
\end{figure}

@@

\subsubsection{Semantic Digital Twin}
From the background information of DTs in Section \ref{subsubsec:DigitalTwins} we created a digital twin, based on a dynamic asset model of a simple building, that handled mobile assets. The asset model had areas with predefined coordinates (latitude1, latitude2, longitude1, and longitude2) The DT and its parts were semantically enriched by using the semantic technologies presented in Technologies, as well as by using SMOL as an enabler of semantics from the analysis of it in Section \ref{subsubsec:SemanticDigitalTwins}.

The \emph{semantic} digital twin was created in SMOL by using its core language illustrated in Section \ref{subsec:SMOL}. And by using the SMOL Interpreter (\verb|smol.jar|) in the file structure to read and execute the SMOL program. The SMOL program consists of two modular parts, namely: 
\begin{enumerate}
    \item \textbf{DS} (Digital Shadow) acts as a timestamped snapshot of a specific smartphone, as well as being used to access the newest sensor data, from the descriptions in Section \ref{subsubsec:FurtherDataSeparation}. It accesses the latest physical locations from a specified time ago. Based on the descriptions in Section \ref{subsec:Databases}, and from the example in Figure \ref{fig:access_time_series}, we accessed the updated sensor data in the DS. This can be seen in the function in Figure \ref{fig:DigitalShadowInflux}. Note that it looks like we get outdated sensor data (\textbf{first} added to the bucket), but we are actually getting the latest. When this was implemented there was no support for single-quote string delimiters in \verb|access|.
    \item \textbf{Digital Twin} is a digital representation of the PT. It accesses: 
    \begin{enumerate}
        \item the critical areas in a building domain defined in the asset model
        \item the smartphones by their IDs from the asset model
        \item the updated physical location of the smartphones from the asset model, based on a revised version of Figure \ref{fig:access_time_series}.
    \end{enumerate}
\end{enumerate}

\begin{figure}[H]
    \centering
    \begin{small}
    \begin{minted}[breaklines]{dart}
List<Double> getSmartphoneCoordinates()
    List<Double> coordinates = null;
    coordinates = access(
        "from(bucket: \"Data\")
            |> range(start: -20m)
            |> filter(fn: (r) => r[\"_measurement\"] == \"data\")
            |> filter(fn: (r) => r[\"_field\"] == \"latitude\" or r[\"_field\"] == \"longitude\")
            |> sort(columns: [\"_time\"], desc: true)
            |> unique(column: \"id\")
            |> aggregateWindow(every: 5m, fn: first, createEmpty: false)
            |> yield(name: \"first\")",
        INFLUXDB("influx.yml")
        );
    return coordinates;
end
    \end{minted}
    \end{small}
    \caption{Accessing the updated sensor data from InfluxDB}
    \label{fig:DigitalShadowInflux}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minted}{dart}
MovableEntity smartphone = new MovableEntity(id, lat, long, False);
    \end{minted}
    \caption{Creating a smartphone object from the smartphone ID in the asset model, in combination with the corresponding updated coordinates from InfluxDB. \textbf{False} simply means the smartphone is not concluded to be endangered yet}
    \label{fig:smartphone_object}
\end{figure}

Referencing the requirement "Check if smartphones are inside any critical areas" we must understand what it \emph{means} before we can implement it. An example is described in Section \ref{subsubsec:RQ2}, where an informed decision can be made if the \emph{criteria} in the example evaluate to be true. The code implementation of it is shown in Figure \ref{fig:inside_critical_area}, where the function is located inside the Area-class. Meaning, we write \verb|area.containsPoint(lat, long)| to check if \verb|area| contains the smartphone with latitude (\verb|lat| and longitude \verb|long|).

\begin{figure}[H]
    \centering
    \begin{small}
    \begin{minted}[breaklines]{dart}
Boolean containsPoint(Double lat, Double long)
    Boolean inAnyLatitudeRange = ((lat >= this.latitude1) & (lat <= this.latitude2)) | ((lat >= this.latitude2) & (lat <= this.latitude1));
    Boolean inAnyLongitudeRange = ((long >= this.longitude1 & long <= this.longitude2)) | ((long >= this.longitude2) & (long <= this.longitude1));

@@

    return False;
end
    \end{minted}
    \end{small}
    \caption{Checking if a smartphone is inside a critical area based on coordinates, assuming they align}
    \label{fig:inside_critical_area}
\end{figure}

If the DT concludes that a smartphone \emph{is} inside \emph{any} of the critical areas which are accessed from the domain, the twin sets \verb|smartphone.endangered = True;| and adds the smartphone object to the specific area's list of endangered smartphones as a result, which is shown in Figure \ref{fig:add_endangered}. Note that what is shown in the figure is inside a while-loop for the context of the counter \textbf{j}.

\begin{figure}[H]
    \centering
    \begin{minted}[breaklines]{dart}
a = areas.get(j);
isInsideACriticalArea = a.containsPoint(lat, long);
if (isInsideACriticalArea) then
    smartphone.endangered = True;
    a.smartphones = new List<MovableEntity>(smartphone, a.smartphones);
end
    \end{minted}
    \caption{Automatically checking if a smartphone is inside a critical area. If the area contains its point (x, y), the smartphone is set to endangered, and is added to the list of the endangered smartphones for the specific area}
    \label{fig:add_endangered}
\end{figure}

Regarding the last requirements of this part, informational messages, such as examples of SPARQL queries, are shown in the REPL during the program execution. Figure \ref{fig:sparql_smol} show how we can \emph{manually} manually check which smartphones are inside a critical and thus are endangered, and the results (information about endangered smartphones). The querying is done in the SMOL REPL, which is started from the information in Section \ref{subsubsec:SMOLInterpreter}.

@@

The DT handles newly added smartphones to the asset model that don't have a physical location as well, and thus all requirements are met.

@@

\section{Evaluation}\label{sec:Evaluation}
This chapter evaluates the developed implementation in terms of which structural requirements in Section \ref{subsec:Requirements} are met from the analysis in this thesis in Chapter \ref{sec:Analysis}, and from performing various test runs in the following sections. We start by presenting the overall evaluation of the implementation according to the requirements. In the following section, we shortly describe how we set up SMOL to be used in the implementation. Then we provide test runs of the server automatically reloading the asset model, and of an operator manually editing the asset model by adding smartphones to it. Then we evaluate how long it takes to warn endangered clients of an informed decision. We also evaluate how the system scales with many heterogeneous smartphones, and plot the results in Data Explorer in InfluxDB.

The Structural requirements were divided into four parts, one for each of the \emph{main} architectural components shown in Figure \ref{fig:components}: \textbf{Client}, \textbf{Server}, \textbf{Databases (\textbf{Firebase Realtime} and \textbf{InfluxDB Time Series}}, and \textbf{Semantic Digital Twin}.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{comment}
\usepackage{duomasterforside}
\usepackage{listings}
\usepackage{float}
\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{fvextra}
\usepackage{soul}
\usepackage{minted}
\AtBeginEnvironment{minted}{%
  \renewcommand{\fcolorbox}[4][]{#4}}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{multirow}
\graphicspath{ {./graphics/} }
\usepackage[backend=biber,natbib,style=numeric]{biblatex}
\usepackage[hidelinks,colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{hyperref}
\addbibresource{references.bib}

@@

\section*{Abstract}
This thesis proposes a solution to the challenge of handling mobile assets in a semantic digital twin (DT) based on a dynamic asset model in which static data (existing building infrastructure) is separated from dynamic data (physical location). The semantic DT easily integrates with semantic technologies, and is enriched by them. A cross-platform app is created for mobile assets to send their status, whilst protecting privacy of data subjects as much as possible. The DT is an up-to-date digital representation of the PT, and it reconfigures in near real-time if the PT changes. An endangered person is warned in the app, concluded to be inside a critical area by the digital counterpart, evaluated to be true for 234 of 234 smartphones. Appropriate clients are warned. We provide informative messages in the SPARQL endpoint provided by the SMOL REPL. The solution is extensible, and can be easily maintained by manual operations. Formalized knowledge of the domain is ensured when automatically reading from it, even if the set of heterogeneous mobile assets change.

@@

Handling mobile assets in a DT based on a dynamic \emph{asset model} (i.e. a file that contains an organized description of an asset's composition and properties \cite{kamburjan_twinning-by-construction_2022}) is unexplored territory.

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.14]{graphics/initial_thesis_overview.png}
    \caption{Overview of \emph{main} architectural components}
    \label{fig:initial_components}
\end{figure}

@@

\subsubsection{For the scope of this thesis}\label{subsubsec:ForTheScope}
As of today, a DT is referred to as a digital representation of some physical system \cite{grieves_digital_2017}, that has no restriction on the size or scale of it, which means the entity could be a single asset or a system of systems for that matter \cite{li_digital_2022, waszak_let_2022}. The physical system exists in real life and is commonly referred to as the physical twin (PT), and is observed by the DT in near real-time. There should be a bidirectional flow of data between the DT and the PT throughout the life cycle of the PT. In other words, the DT gets updated sensor data from the PT, and sends informed decisions back to it \cite{madni_leveraging_2019, waszak_let_2022, kamburjan_digital_2022}. From this, and for the scope of this thesis, a DT is a digital representation of a system that tracks many mobile assets. The DT sends informed decisions to the physical counterpart in which the appropriate mobile assets perform actions.

There are multiple misconceptions about DTs. One such misconception is that a digital shadow (DS) is a DT. This misconception often leads people to build shadows instead of DTs. \cite{fuller_digital_2020, li_digital_2022}.  Although very similar, a digital shadow (DS) does not have the bidirectional data flow that a DT does. Instead, there is a one-way flow from the PT to the digital counterpart, which often leads to a change in the DS, but not vice versa \cite{kritzinger_digital_2018, li_digital_2022}. Secondly, as told by \citeauthor{fuller_digital_2020}, a common misconception is that the DT must be an exact 3D copy of a physical entity, such as a building.

@@

To sum up, it is important to understand what a DT is and what it is not. We should try to "close the loop" by sending informed decisions back to the physical counterpart so that it reflects the newest state of the DT. A DS could be used alongside the DT with the only goal of being a timestamped snapshot of some specific physical asset \cite{bergs_concept_2021}. When implementing one should also think about the challenges of digital twins. We should not get too caught up with modeling an exact 3D replica of an asset. On the other hand, we \emph{should} use existing semantic technologies, such as knowledge graphs, to formally represent knowledge from a variety of data sources.

@@

OWL documents can easily be created (from exporting), loaded into, and maintained in ontology editors, such as Protégé\footnote{\url{https://protege.stanford.edu}}. Such a tool can be used to give semantics to e.g. a model of a building, as well as spending less time on creating and maintaining it compared to BIM.

@@

\begin{figure}[H]
    \centering
    \caption{A simple SMOL program (example.smol)}
    \label{fig:smol_program}
    \begin{Verbatim}[frame=single]
class Smartphone(Int id, String name, Boolean isBlackBerry) end

@@

\subsection{Reasoning}\label{subsec:Reasoning}
OWL was previously mentioned in Section \ref{subsubsec:Technologies} as a language for expressing rich knowledge of concepts and relations between entities. We should keep in mind, however, that this formalized knowledge should be reasoned over and put in context to draw conclusions based on criteria. There are many technologies that let us do this.
\subsubsection{HermiT}
HermiT is a reasoner for OWL 2 \cite{glimm_hermit_2014}, which is an upgraded version of OWL. \citeauthor{glimm_hermit_2014} describes that the reasoner support both object and data property classification, as well as SPARQL query answering. HermiT is built into the ontology editor Protégé and can be used to reason over an ontology directly in the editor. Other plugins for reasoning can be added to the ontology editor, but it is convenient that HermiT comes pre-installed with the desktop version.

@@

There also exist other technologies that provide reasoner tools. Some notable mentions include, but are not limited to:
\begin{itemize}
    \item{\textbf{Apache Jena}} is an open source framework that has the Inference API which include reasoner tools that can be used to reason over data, as well as checking content of triple stores \cite{noauthor_apache_nodate-1}. The source code is publicly available\footnote{\url{https://github.com/apache/jena}}.
    \item{\textbf{OWL API}} is a Java framework that lets us create, serialize or manipulate ontologies written in OWL \cite{noauthor_owl_nodate}, and its source code is openly accessible\footnote{\url{https://github.com/owlcs/owlapi}}. Its recent development is more focused towards OWL 2 however.
    \item{\textbf{SMOL}} provide reasoning tools through its interpreter, which implements \emph{semantic lifting} (i.e. generating a knowledge graph from the current state of a SMOL program) \cite{noauthor_semantic_nodate}.

@@

\subsection{Research questions (RQ1, RQ2, RQ3)}
\subsubsection{RQ1}\label{subsubsec:RQ1}
RQ1 is concerned with creating a dynamic asset model of a simple building. From the descriptions of asset models in Section \ref{subsec:AssetModel}, an ontology of a simple building was created, as shown in Figure \ref{fig:building_ontology}. This asset model is a simplified version of an exported OWL document (file). It can be created in an ontology editor, such as Protégé. Although other ontology editors can be used, we should use Protégé due to its user-friendliness and the extensive documentation provided in its user guide \cite{horridge_practical_2011}. This exported file is provided to and used by the digital twin (DT) by using the SMOL interpreter. Section \ref{subsubsec:DynamicAssetModel} further examines what a \emph{dynamic} asset model entails.

@@

\begin{small}
\begin{Verbatim}[commandchars=\\\{\}, breakanywhere=true]
### \hl{http://www.semanticweb.org/oscarlr/ontologies/2023/2/building#smartphone5}
\hl{:smartphone5} rdf:type owl:NamedIndividual ,
                      :MovableEntity ;
             :movableEntityId "\hl{5}" .
\end{Verbatim}
\end{small}

@@

\subsubsection{RQ3}
RQ3 is concerned with how we can "separate dynamic data (smartphone’s physical location) from static data
(building infrastructure) in the asset model". Hypothesis (H) states that the asset model should be of a simple building. The rooms with a square area constitute it. In contrast to building infrastructure, mobile assets that are tracked can change at any time, as well as their continuously changing physical locations. As described in Section \ref{subsec:Scope}, we are not concerned with other movable entities in the building domain apart from the mobile assets. And, even though the data of the building infrastructure may change, it will not change during program execution in SMOL. This is not the case with the time series data accessed from InfluxDB in SMOL. However, one way to separate the dynamic data from the static data is by examining \emph{how} often changes occur through time, where separate means not affecting one another. More specifically, when reloading the asset model, the static data should not be affected by the data data considered dynamic. Further analysis of static and dynamic data is described in Section \ref{subsubsec:StaticAndDynamicData}, and also later discussed in Section \ref{subsec:DiscussDynamicData}.
\newline

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.14]{graphics/static_built_environment.png}
    \caption{A UML class diagram that illustrates the separation of static and dynamic behavior}
    \label{fig:static_built_environment}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.18]{graphics/static_dynamic_asset_model.png}
    \caption{The separation of static and dynamic data illustrated in blue and red, respectively}
    \label{fig:static_dynamic_asset_model}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.12]{graphics/wireframe_app.png}
    \caption{Wireframe of the UI in the app}
    \label{fig:wireframe_app}
\end{figure}

@@

Then we look at the final version in Figure \ref{fig:components}, and describe what was implemented. For each of the larger architectural components in the figure, we reference the section that describes it in Chapter \ref{sec:Background}, and the design decision that enables it in Chapter \ref{sec:Analysis}. We also reference the corresponding \emph{item} in the structural requirement \emph{part} we are in. Along the way, we discuss challenges when developing, as well as which requirements were met and why. Due to this, the last section presents the overall evaluation of the implementation in terms of the structural requirements. Further evaluation is presented in Chapter \ref{sec:Evaluation}.

@@

An asset model (ontology) of a simple building was also created, as shown in Figure \ref{fig:simple_building} (this time with coordinates as data properties which were "added" to individuals of the \verb|MovableEntity| class suggested in Section \ref{subsubsec:DynamicAssetModel}. We used Protégé to create the domain and exported it as an OWL document (file). This file was then accessed in a new SMOL program based on the background information in Section \ref{subsubsec:Technologies}, Section \ref{subsec:SMOL}, and from further analysis in Section \ref{subsubsec:StaticAndDynamicData}.

@@

Figure \ref{fig:server_forward_informed_decision} shows how the informed decision from the DT is forwarded to Firebase.
\begin{figure}[H]
    \centering
    \begin{small}
    \begin{minted}[breaklines]{java}
private void warnEndangeredClients(Map<String, String> results) {
    reference = FirebaseDatabase.getInstance().getReference("endangered");
    DatabaseReference.CompletionListener completionListener = (databaseError, databaseReference) -> {
        System.out.println("Forwarded informed decision to PT!");
        hasWarned = true;
    };
    reference.setValue(results, completionListener);
}
    \end{minted}
    \end{small}
    \caption{Forward informed decision to PT through Firebase}
    \label{fig:server_forward_informed_decision}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{small}
    \begin{minted}{java}
public void insertDataPoint(Object object) {
    WriteApiBlocking writeApi = db.getWriteApiBlocking();
    writeApi.writeMeasurement(bucket, org, WritePrecision.MS, object);
}
    \end{minted}
    \end{small}
    \caption{Inserting a measurement point (timestamped status message) into the Data-bucket in InfluxDB}
    \label{fig:insert_point}
\end{figure}

@@

\subsubsection{Semantic Digital Twin}
From the background information of DTs in Section \ref{subsec:DigitalTwins} we created a digital twin, based on a dynamic asset model of a simple building, that handled mobile assets. The asset model had areas with predefined coordinates (latitude1, latitude2, longitude1, and longitude2) The DT and its parts were semantically enriched by using the semantic technologies presented in Technologies, as well as by using SMOL as an enabler of semantics from the analysis of it in Section \ref{subsubsec:SemanticDigitalTwins}.

The \emph{semantic} digital twin was created in SMOL by using its core language illustrated in Section \ref{subsec:SMOL}. And by using the SMOL Interpreter (\verb|smol.jar|) in the file structure to read and execute the SMOL program. The SMOL program consists of two modular parts, namely: 
\begin{enumerate}
    \item \textbf{DS} (Digital Shadow) has to purposes. First, it acts as a timestamped snapshot of smartphones at a certain point in time. From the background information in Section \ref{subsubsec:ForTheScope}, we abstract this so that the asset is all tracked mobile assets. Secondly, it is used to access the newest sensor data, from the descriptions in Section \ref{subsubsec:FurtherDataSeparation}. It accesses the latest physical locations from a specified time ago and only gets sensor data. This is based on the descriptions in Section \ref{subsec:Databases}, and from the example in Figure \ref{fig:access_time_series}. This can be seen in the function in Figure \ref{fig:DigitalShadowInflux}. Note that it looks like we get outdated sensor data (\textbf{first} added to the bucket), but we are actually getting the latest. Also note that when this was implemented there was no support for single-quote string delimiters in \verb|access|.
    \item \textbf{Digital Twin} is a digital representation of the PT. It accesses: 
    \begin{enumerate}
        \item the critical areas in a building domain defined in the asset model
        \item the smartphones by their IDs from the asset model
        \item the updated physical location of the smartphones from the asset model, based on a revised version of Figure \ref{fig:access_time_series}.
    \end{enumerate}
\end{enumerate}

\begin{figure}[H]
    \centering
    \begin{small}
    \begin{Verbatim}[frame=single,breaklines]
List<Double> getSmartphoneCoordinates()
    List<Double> coordinates = null;
    coordinates = access(
        "from(bucket: \"Data\")
            |> range(start: -20m)
            |> filter(fn: (r) => r[\"_measurement\"] == \"data\")
            |> filter(fn: (r) => r[\"_field\"] == \"latitude\" or r[\"_field\"] == \"longitude\")
            |> sort(columns: [\"_time\"], desc: true)
            |> unique(column: \"id\")
            |> aggregateWindow(every: 5m, fn: first, createEmpty: false)
            |> yield(name: \"first\")",
        INFLUXDB("influx.yml")
        );
    return coordinates;
end
    \end{Verbatim}
    \end{small}
    \caption{Accessing the updated sensor data from InfluxDB}
    \label{fig:DigitalShadowInflux}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{minted}{dart}
MovableEntity smartphone = new MovableEntity(id, lat, long, False);
    \end{minted}
    \caption{Creating a smartphone object from the smartphone ID in the asset model, in combination with the corresponding updated coordinates from InfluxDB}
    \label{fig:smartphone_object}
\end{figure}

As described in Section \ref{subsec:AssetModel}, a DT should allow for interoperability between various sources of data. By using the common class \verb|MovableEntity| for the entities in the set of movable heterogeneous entities, interoperability was gained in the DT.

Referencing the requirement "Check if smartphones are inside any critical areas" we must understand what it \emph{means} before we can implement it. An example is described in Section \ref{subsubsec:RQ2}, where an informed decision can be made if the \emph{criteria} evaluate to be true. The code implementation of it is shown in Figure \ref{fig:inside_critical_area}, where the function is located inside the Area-class. Meaning, we write \verb|area.containsPoint(lat, long)| to check if \verb|area| contains the smartphone with latitude (\verb|lat|) and longitude (\verb|long|).

\begin{figure}[H]
    \centering
    \begin{small}
    \begin{Verbatim}[frame=single,breaklines]
Boolean containsPoint(Double lat, Double long)
    Boolean inAnyLatitudeRange = ((lat >= this.latitude1) & (lat <= this.latitude2)) | ((lat >= this.latitude2) & (lat <= this.latitude1));
    Boolean inAnyLongitudeRange = ((long >= this.longitude1 & long <= this.longitude2)) | ((long >= this.longitude2) & (long <= this.longitude1));

@@

    return False;
end
    \end{Verbatim}
    \end{small}
    \caption{Checking if a smartphone is inside a critical area based on coordinates, assuming they align}
    \label{fig:inside_critical_area}
\end{figure}

If the DT concludes that a smartphone \emph{is} inside \emph{any} of the critical areas which are accessed from the domain, the twin sets \verb|smartphone.endangered = True;| and adds the object to the list of endangered smartphones for the specific area, which is shown in Figure \ref{fig:add_endangered}. Note that what is shown in the figure is inside a while-loop, hence the use of the counter \textbf{j}.

\begin{figure}[H]
    \centering
    \begin{Verbatim}[frame=single,breaklines]
a = areas.get(j);
isInsideACriticalArea = a.containsPoint(lat, long);
if (isInsideACriticalArea) then
    smartphone.endangered = True;
    a.smartphones = new List<MovableEntity>(smartphone, a.smartphones);
end
    \end{Verbatim}
    \caption{Checking if a smartphone is inside a critical area. If the area contains its point (x, y), the smartphone is set to endangered, and is added to the list of the endangered smartphones for the specific area}
    \label{fig:add_endangered}
\end{figure}

Regarding the last requirements of this part, informational messages, such as examples of SPARQL queries, are shown in the REPL during the program execution. Figure \ref{fig:sparql_smol} show how we can \emph{manually} check which smartphones are inside a critical and thus are endangered, and the results (information about endangered smartphones). The querying is done in the SMOL REPL, which is started from the information in Section \ref{subsubsec:SMOLInterpreter}.

@@

The DT handles newly added smartphones to the asset model that don't have a physical location as well. Meaning the knowledge graph reflect there are smartphones in the domain without a physical location, but the system does not however track them. This means the the requirement "Handle new smartphones that don't have a physical location" is not \emph{fully} met.

@@

\subsection{Which requirements are met?}\label{subsec:WhichRequirements}
This section \emph{presents} the overall evaluation of the developed implementation in terms of which structural requirements in Section \ref{subsec:Requirements} were met, based on the analysis in Chapter \ref{sec:Analysis}. The Structural requirements were divided into four parts, one for each of the \emph{main} architectural components shown in Figure \ref{fig:components}: \textbf{Client}, \textbf{Server}, \textbf{Databases (\textbf{Firebase Realtime} and \textbf{InfluxDB Time Series}}, and \textbf{Semantic Digital Twin}. Table \ref{tab:overall_evaluation} \emph{presents} the overall evaluation.
